Stack Safety is a Security Property

### 2021-04-11 tex/main.tex
More discussion of properties relating to the broader world

\bcp{If we end up adopting my recent proposal for defining subtraces and
  separating WBCF, I think we can be a bit stronger here: We define stack
  safety in IFC terminology based on the same intuitions as their LSE, but
  we consider (and define) WBCF to be a separate thing---desirable, but
  distinct from safety.  (If we make this argument, we'll need to show an
  example of a program that is stack safe but doesn't have WBCF.)}In this
work, we take the (informal) definitions of WBCF and LSE as a starting
point, but choose a different approach to formalizing them, based on the
technical framework of language-based security~\citep{??}. We define
trace-based properties of \emph{stepwise stack integrity} and \emph{stack confidentiality}.
In applying these security concepts, we treat each caller-callee pair as
mutually distrustful, with the caller needing protection from the callee. \apt{Need more explanation of
  ``nesting'' and our novelty factor here? Also about domains??}
This differs from common instances in language-based security in that the
confidentiality property applies to many nested subtraces within a single execution.
\sna{Thinking about how to express domains, or whether to. I'm not sure it's
  an interesting detail at this level. Also not sure how strongly to talk about this novelty.}
{\em Integrity} says that a caller's local data is protected from modification by its callee.
\bcp{What about the ``stepwise'' part?}
This is vital for local reasoning about the behavior of the
caller.\bcp{Either delete that sentence or explain it more.}
{\em Confidentiality} says that a caller's data is protected
from being leaked into its callee, and also that the callee cannot see uninitialized data.
As a reasoning principle, confidentiality guarantees that the callee's
behavior is determined only by a predictable interface that does not include the caller's
local state or the uninitialized portion of the stack.
Together, these properties (informally) capture LSE, as well as the ``temporal stack safety''
notion that data outside of live frames should be inaccessible (which LSE
does not address\bcp{Which supports my feeling that we should not advertise
  ourselves as reimplementing LSE but rather as defining a different
  condition based on similar intuitions.}).
\bcp{These are the intuitions, but we should also mention that we want to define
  these (at least partly) in terms of a sort-of-standard notion of ``nested
  noninterference''!  Then, at the end of the intro (just before the
  contributions) we can summarize the whole thing like ``So what is stack
  safety?  It is nested NI.''}
We also give a simple trace-based formalization of WBCF, which is essential for keeping the
notions of ``caller'' and ``callee'' sensible. \bcp{That last should probably
be a separate discussion, if I've succeeded in convincing people that WBCF
is not part of stack safety.}

For example, \citet{DBLP:conf/sp/RoesslerD18} describe how to protect the stack
using tag-based \emph{micro-policies}~\citep{pump_oakland2015}.
Intuitively, memory cells within each stack frame are tagged with the
identity of the function activation
that owns the frame, and access to those locations is then gated
according to the activation that is currently executing. Roessler and Dehon
explore alternative versions of this basic idea, with different ways of identifying
functions and deciding exactly which accesses are allowed, to enable different ways
of maintaining the connection between an activation identity and the memory it owns.
These vary in their efficiency and in the spatial and temporal precision of their security
guarantees.
For example, their \emph{Depth Isolation} policy  requires a strict mapping from a function
to its frame from the start of the call to its end, and permits the function to write
only into that frame.
This seems simple and intuitively correct, but it can be expensive to implement the strict
mapping. Conversely, their \emph{Lazy Tagging and Clearing} policy improves
performance by assigning locations to the active function when they are written, which
permits  ``illicit'' writes to inaccessible stack frames so long as the
written locations are not subsequently read by their ``real'' owners. The justification
for this form of laziness is that any genuinely dangerous writes will ultimately be
detected before they can affect program behavior.

To illustrate the applicability of these properties, we perform randomized testing to
validate that \citeauthor{DBLP:conf/sp/RoesslerD18}'s Depth Isolation micro-policy enforces
them. Randomized testing is a relatively low-cost way to validate a potential enforcement
mechanism; it finds flaws in the enforcement fast enough to be integrated into the workflow
of enforcement development \citep{TestingNI:ICFP}. To demonstrate that the testing adequately
covers potentially erroneous programs, we test intentionally buggy policies alongside
the real one. If the testing can differentiate correct implementations from buggy ones,
it is an argument in favor of their correctness.
%\apt{more to say?}\bcp{Yes---partly to underscore our claim that this is a
%  significant contribution, and partly to give the reader some context for
%  why this is part of our story.  E.g., ``One nice thing we can do with our
%  spec is to use it to evaluate the claims of existing or future proposals
%  for enforcement mechanisms.  One relatively inexpensive way to carry out
%  such an evaluation is to see whether property-based testing against this
%  spec can distinguish correct implementations of the mechanism from buggy
%  ones...''}

However, as our testing reveals, \citeauthor{DBLP:conf/sp/RoesslerD18}'s Lazy Tagging and
Clearing micro-policy does not enforce stepwise integrity.
To address this, we develop an \emph{observational stack integrity} property that
is intended to specify enforcement mechanisms that
detect integrity violations between the point when they occur and the point when they make
an observable impact. Intuitively this property can be phrased as ``no visible reads after dangerous writes.''  We prove (in Coq) that stepwise integrity implies observational integrity.
\sna{There is a bit of massaging old proofs to do here, and I don't recall if we
  established that this was important.}

Unfortunately the existing Lazy Tagging and Clearing
micro-policy does not quite enforce this property either, as we show in
testing~\apt{really??}, but formalizing the ideal property allows us to see how it falls short.
We propose  a variant of Lazy Tagging and Clearing that we believe \emph{does} enforce
observational integrity, at some performance cost.\apt{but we're not testing that, right?}
\rb{Not sure about the property part right now, the machine side of this is
  pretty much done}
\sna{It's aspirational so far, but there's no fundamental reason we can't test it. We may
  need a caveat about how it's harder to test?}

Observational integrity is attractive in being significantly more \emph{extensional} than
stepwise integrity.  We therefore argue that, despite its somewhat greater complexity,
it may be a better minimal specification for
stack safety enforcement mechanisms in general.

@@

Intuititively, stack safety means that a caller's stack frame is protected
from a confused or malicious callee's
attempts to read or write its private data or hijack its control flow. And in fact the caller,
while presumably\bcp{not ``presumably'': {\em by definition}!} not malicious, may itself be confused or erroneous. Our attacker
model reflects this intention\bcp{I don't understand how it reflects this
  intuition about the caller---indeed, I don't understand the point about
  the caller.} by assuming that both the attacker and the defender
may run arbitrary machine code. Hardware and timing attacks are out of scope.

>>>>

\bcp{If we end up adopting my recent proposal for defining subtraces and
  separating WBCF, I think we can be a bit stronger here: We define stack
  safety in IFC terminology based on the same intuitions as their LSE, but
  we consider (and define) WBCF to be a separate thing---desirable, but
  distinct from safety.  (If we make this argument, we'll need to show an
  example of a program that is stack safe but doesn't have WBCF.)}
In this work, we take the (informal) definitions of WBCF and LSE as a starting
point, but choose a different approach to formalizing them, based on the
technical framework of language-based security~\citep{??}. We define
trace-based properties of \emph{stepwise stack integrity} and
\emph{lockstep stack confidentiality}.
In applying these security concepts, we treat each caller-callee pair as
mutually distrustful, with the caller needing protection from the callee. \apt{Need more explanation of
  ``nesting'' and our novelty factor here? Also about domains??}
This differs from common instances in language-based security in that the
confidentiality property applies to many nested subtraces within a single execution.
\sna{Thinking about how to express domains, or whether to. I'm not sure it's
  an interesting detail at this level. Also not sure how strongly to talk about this novelty.}
{\em Integrity} says that a caller's local data is protected from modification by its callee,
and its {\em stepwise} interpretation formalizes this very naturally as a caller's local data
never being modified during a call. {\em Confidentiality} says that a caller's data is protected
from being leaked into its callee, and also that the callee cannot see uninitialized data.
As a reasoning principle, confidentiality guarantees that the callee's
behavior is determined only by a predictable interface that does not include the caller's
local state or the uninitialized portion of the stack. This is formalized as a
{\em noninterference}-style property that, rather than comparing pairs of whole program traces,
compares a pair of subtraces for each call, nested within other calls. The term
{\em lockstep} describes the behavior of these pairs: they step together, and any
deviation is recognized as a violation of confidentiality.
Together, these properties capture the rough intuition of LSE, but are not identical --
confidentiality contains temporal protections that LSE does not address.

Since our stack safety occupies the same conceptual space as LSE, we should consider how it
relates to WBCF. First, we observe that some degree of bracketed control flow is implicit
in any notion stack safety, simply because calls and returns must be recognizeable; callers
and callees, distinct. We treat control flow that creates an ambiguity between caller and
callee as keeping the callee in control, and therefore such behavior can never give access
to the caller's data. An example would be if the callee attempts to ``return'' to the
wrong location, hoping to execute code on behalf of the caller: it may make the jump
without violating stack safety, but it gains no privileges on the caller's data.
Alternatively a callee might return past its immediate caller, to somewhere deeper in the
stack, and stack safety does not inherently rule this out. We then define WBCF as an additional
restriction, describing the standard control flow principle that a return pops exactly one frame
from the stack.
\sna{We got side-tracked a bit from talk of entry integrity, and Andrew made some good points
  about it, so let's not rule that out just yet. But I focused on returns because there's a
  clear story already.}

Part of the challenge in fully characterizing stack safety is that at a low level,
without built-in stacks, we are at the mercy of architectural choices that
do not relate to the abstract notion of stack safety. We abstract over these choices
as much as possible, so that our model can be applied to a wide range of architectures
and enforcement mechanisms. But a second challenge is the diversity of enforcement
mechanisms and their need for efficiency. In the name of efficiency, enforcement
mechanisms may weaken their protections, arguing that the protections they do provide
still capture stack safety in a meaningful sense. This calls for an {\em extensional}
characterization of stack safety, one that describes the externally visible behavior
of a stack safe system. Our stepwise and lockstep properties are {\em intensional},
focusing on the internal behavior.

For example, \citet{DBLP:conf/sp/RoesslerD18} describe how to protect the stack
using tag-based \emph{micro-policies}~\citep{pump_oakland2015}.
Intuitively, memory cells within each stack frame are tagged with the
identity of the function activation
that owns the frame, and access to those locations is then gated
according to the activation that is currently executing. Roessler and Dehon
explore alternative versions of this basic idea, with different ways of identifying
functions and deciding exactly which accesses are allowed, to enable different ways
of maintaining the connection between an activation identity and the memory it owns.
These vary in their efficiency and in the spatial and temporal precision of their
security guarantees.
Their \emph{Depth Isolation} policy requires a strict mapping from a function
to its frame from the start of the call to its end, and permits the function to write
only into that frame.
This seems simple and intuitively correct, but it can be expensive to implement the strict
mapping. Conversely, their \emph{Lazy Tagging and Clearing} policy improves
performance by assigning locations to the active function when they are written, which
permits  ``illicit'' writes to inaccessible stack frames so long as the
written locations are not subsequently read by their ``real'' owners. The justification
for this form of laziness is that any genuinely dangerous writes will ultimately be
detected before they can affect program behavior.

To address this, we develop an \emph{observational integrity} property that
is supports enforcement mechanisms that detect violations between the point when
they occur and the point when they make an observable impact. Intuitively this property
can be phrased as ``no visible reads after dangerous writes.''  We prove (in Coq) that stepwise
integrity implies observational integrity.
\sna{There is a bit of massaging old proofs to do here, and I don't recall if we
  established that this was important.}
We additionally sketch an associated {\em observational confidentiality} property
makes confidentiality extensional in the same way, though we do not know of a real
enforcement mechanism that weakens confidentiality in this way. Together,
these form an {\em observational stack safety} that is highly extensional.
We therefore argue that, despite its somewhat greater complexity,
it may be a better minimal specification for
stack safety enforcement mechanisms in general.

To illustrate the applicability of these properties, we perform randomized testing to
validate that \citeauthor{DBLP:conf/sp/RoesslerD18}'s Depth Isolation micro-policy enforces
them. Randomized testing is a relatively low-cost way to validate a potential enforcement
mechanism; it finds flaws in the enforcement fast enough to be integrated into the workflow
of enforcement development \citep{TestingNI:ICFP}. To demonstrate that the testing adequately
covers potentially erroneous programs, we test intentionally buggy policies alongside
the real one. If the testing can differentiate correct implementations from buggy ones,
it is an argument in favor of their correctness.
%\apt{more to say?}\bcp{Yes---partly to underscore our claim that this is a
%  significant contribution, and partly to give the reader some context for
%  why this is part of our story.  E.g., ``One nice thing we can do with our
%  spec is to use it to evaluate the claims of existing or future proposals
%  for enforcement mechanisms.  One relatively inexpensive way to carry out
%  such an evaluation is to see whether property-based testing against this
%  spec can distinguish correct implementations of the mechanism from buggy
%  ones...''}
Of course, as our testing reveals, \citeauthor{DBLP:conf/sp/RoesslerD18}'s Lazy Tagging and
Clearing micro-policy does not enforce stepwise integrity.
Unfortunately the existing Lazy Tagging and Clearing
micro-policy does not quite enforce observable confidentiality either. %, as we show in
%testing~\apt{really??},
but formalizing the ideal property allows us to see how it falls short.
We propose  a variant of Lazy Tagging and Clearing that we believe \emph{does} enforce
observational integrity, at some performance cost.\apt{but we're not testing that, right?}
\rb{Not sure about the property part right now, the machine side of this is
  pretty much done}
\sna{It's aspirational so far, but there's no fundamental reason we can't test it. We may
  need a caveat about how it's harder to test?}

@@

Intuititively, stack safety means that a caller's stack frame is protected
from a confused or malicious callee's
attempts to read or write its private data or hijack its control flow. And in fact the caller,
while not malicious, may itself be confused or erroneous. Our attacker
model reflects this intention\bcp{I don't understand how it reflects this
  intuition about the caller---indeed, I don't understand the point about
  the caller.} by assuming that both the attacker and the defender
may run arbitrary machine code. Hardware and timing attacks are out of scope.
